---
title: "index"
author: "Israel Martinez"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(knitr)
library(ggplot2)
library(leaflet)
source("analysis.R")
```

## Summary Information
The data reveals that the amount of shooting events that occured in 2018 was
**`r shooting_events`**. The amount of lives that were lost in 2018 due to
shootings was **`r lives_lost`**. In regard to the city that was most impacted
by shootings (cities with the highest number of killed and injured shootings),
the city was **`r cty_most_impact`**. For my two other insights, I wanted to
know the month with the most shootings: **`r month_most_shoot`**. I also wanted
to know which state had the highest percentage of kills in their shootings: 
**`r state_kill_rate`**.

This information can be found [here](http://www.shootingtracker.com/).

## Summary Table
```{r table, echo=FALSE}
table_caption <- ("The state that had the most casualties was California. The
                  casualties are a combination of those killed and those
                  injured.")
kable(summary_table, caption = table_caption)
```

## Description of a particular incident
The incident that I have chosen to describe is the event which had the most
number of shootings (killed and injured). The data revealed that the event with
the highest number was at **`r worst_incident$address`,**
**`r worst_incident$city`, `r worst_incident$state`**. This event occured on
**`r worst_incident$date`** and there were **`r worst_incident$casualties`**
casualties (**`r worst_incident$num_killed`** killed and
**`r worst_incident$num_injured`** injured). You can read a personal Time's
article piece about this shooting incident
[here](https://time.com/parkland-high-school-shooting-vigil/).

## An interactive map
The purpose of this interactive map is to showcase where the shooting incidents
occured in the United States. Seeing the geographic location itself is
important because it can reveal certain trends.
```{r map, echo=FALSE}
leaflet(data = more_column) %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(lng = -97.922211, lat = 39.381266, zoom = 4) %>%
  addCircleMarkers(
    lng = ~long,
    lat = ~lat,
    popup = ~paste("Address:", address, "<br>",
                   "Date:", date, "<br>",
                   "Casualties:", num_killed, "killed,", num_injured,
                   "injured"),
    stroke = FALSE
  )
```
According to this map, it looks like most of the shootings in 2018 occured in
the Southeast region of the United States. Also, there shows to be zero
shootings in the Northern states such as Montana, Idaho, and North Dakota.

## Plot of your choice
For this plot, I wanted to know the relationships between months and number of casualties per month so I can see any trends. So, I chose to use a bar graph to visualize the data.

```{r graph, echo=FALSE}

# I couldn't figure out how to get the x axis of month to be in chronological
# order, so I used this solution:
# https://thepracticalr.wordpress.com/tag/ggplot2/
more_column$month <- factor(more_column$month,
                       levels = c("January", "February", "March", "April",
                                  "May", "June", "July", "August",
                                  "September", "October", "November",
                                  "December"))

ggplot(more_column) +
  geom_col(mapping = aes(x = month, y = casualties)) +
  labs(
    title = "2018 Shooting Casualties"
  )
```

Looking at the graph, I first notice that the top 3 months with the highest
casualties were summer months. Also, as the year went by after summer, the
number of casualties slowly decreased overtime.

## Critical Reflection
I believe that this report can be used in many good ways. For example,
legislators who write laws for gun control can use the map and maybe direct
their energy toward states with the most shooting incidents (the Southeast for
instance). However, if someone were to just focus on the map and work on areas
with the most shootings, they most likely wouldn’t address the Chicago
community which is the city with the highest number of casualties in the year.
A big group of people could be neglected because of this.